Assessing the Accessibility of Protein Tunnels using Motion Planning

### 2017-12-09 paper.tex
first 2 pages checked

\begin{abstract}
Chemical interactions between proteins and other molecules (ligands) take place in specific locations, called active sites, that are often deeply buried inside the protein structure.
These active sites are accessible through one or more void paths, called tunnels.
Nowadays, tunnels are mostly computed using Voronoi diagrams which approximate the ligand by a bounding sphere.
This leads to highly imprecise results.
To address this problem, we define the computation of the transportation path of non-spherical ligands including their conformation changes (i.e., when the atoms of the ligand move relatively to each other) as a motion planning task in high-dimensional configuration space.
This can be solved using sampling-based planning methods.
We propose a modification of the Rapidly Exploring Random Tree planner for the purpose of trajectory computations of flexible ligands.
To cope with the narrow passage problem, the random samples are generated around a Voronoi-based tunnel.
%which serves us for better localization of the searched space.
The flexibility of the ligand is modeled using a fixed set of known conformations, which does not increase the number of dimensions of the configuration space.
The proposed method has been evaluated in a virtual screening scenario to estimate the traversability of tunnels for several ligands.
\end{abstract}

@@

Reactivity of proteins with other molecules plays a crucial role in many research disciplines, including protein engineering and drug design.
Protein reactivity is tightly connected with the presence of the void space between its amino acids. 
This space, forming a path called tunnel, can be utilized by a small molecule (ligand) which can enter the protein inner space and react with its amino acids surrounding a specific site, called the active site~\cite{gora2013gates}.
\textcolor{blue}{The size and location of tunnels determine the ligand size and shape which can enter the active site.}
Currently, the biochemists use various computational methods for tunnel detection (Fig.~\ref{fig::motiv}) in order to estimate the possibility of interactions between a protein and a ligand.
This helps them to reduce the number of necessary in-vitro lab experiments.
The most common tools detect the tunnels using Voronoi diagrams~\cite{yaffe2008,caver3} assuming a spherical ligand (probe), 
i.e., they approximate the ligand by a bounding sphere.
However, the ligands are typically of a non-spherical shape which is completely ignored by these methods.

@@

Our approach aims to overcome this problem by taking into account the actual shape of the ligand.
This necessarily requires to consider also conformational changes, i.e., the mutual position of the ligand atoms can change.
Such a ligand is denoted as flexible ligand.
%Our approach aims to overcome this problem by taking into account the actual shape of the ligand and also its conformations.
%\textcolor{blue}{To provide the biochemists with a solution which detects tunnels also in these frames and understand the ligand passage, it is necessary to take into account not only the ligand shape but also its conformation changes when the ligand moves.}
%The proposed solution needs also to cope with the narrow passage problem, as proteins are dense structures limiting the ligand movements inside.
%Such a ligand, for which the mutual position of its atoms can change, is denoted as a flexible ligand.

The analysis of traversability of the tunnels by flexible ligands can be formulated as a motion planning problem and
solved by sampling-based planners.
Sampling-based planners randomly sample the configuration space of the robot (ligand) and connect the collision-free samples
into a graph structure (roadmap).
Rapidly Exploring Random Tree (RRT)~\cite{lavalleRRT} is one of the most used sampling-based planners.
Besides many applications in robotics, the sampling-based planners have been already used in biochemistry as well, 
e.g., to study loop motions~\cite{cortes2004geometric}, to identify binding sites~\cite{bayazit2001ligand}, for exit
pathway detection~\cite{cortes2010simulating,cortes2005path} and especially for protein folding~\cite{raveh2009rapid,amato2002using}. %novinskaya2015improving

@@

In this paper, we propose a modification of the RRT planner to find trajectories for non-spherical flexible ligands inside protein tunnels.
The random samples are generated around a given tunnel instead in the whole configuration space.
%We localize the generation of the random samples, i.e., instead of sampling the whole configuration space we focus only on the space surrounding a potential tunnel detected by the Voronoi-based method (using a smaller ligand probe in order to avoid omitting possibly feasible tunnel candidates).
The potential tunnel is represented by a set of consecutive spheres surrounding the tunnel centerline.
To enable the movement of the ligand in the narrow parts of the tunnel, the free-space is dilated by shrinking the atom radii, similarly, e.g., to~\cite{hsu06multilevel}.  %cortes2005path removed as it is not clear it they really rescale
The ligand flexibility is modeled using a fixed set of predefined conformations which the ligand can occupy.

Finding trajectories only along a given tunnel tunnel is faster than computing any exit pathway from the protein due to decreased
volume of the configuration space to be searched.
Another advantage of our method in comparison with the existing approaches is that we enable the users to trace the ligand path leading from the
\textcolor{blue}{outside/outer?} solvent towards the protein active site.

The proposed method was tested by the biochemists who compared the results with already known trajectories of ligands, determined by their traditional approaches.
In the following section, we first mention several approaches and methods related to our solution and then we describe the proposed method in detail.

@@

Most of the available tools for tunnel detection are based on Voronoi diagrams~\cite{yaffe2008,caver3} or they utilize a grid-based computation~\cite{sehnal2013mole}.
%The tunnels are pathways computed for a single atom (spherical probe) using Voronio-diagrams~\cite{yaffe2008,caver3} or 
%grid-based approaches~\cite{sehnal2013mole,petrek2006caver}.
The tunnels are represented as a sequence of spheres and characterized by their bottleneck, length, curvature, and a list of surrounding residues, whose physico-chemical properties are crucial for assessing the interaction possibilities between the protein and ligand.
More details can be found in the recently published state-of-the-art report about analysis and visualization of biomolecular cavities~\cite{Krone_2016}.
The main disadvantage of both grid-based and VD-based methods is that the shape of the ligand is not taken into account during the tunnel detection and it is therefore not easy to estimate if (and how) a non-spherical ligand can traverse the tunnel.
The ligand flexibility cannot be principally considered by these approaches.
Simulations based on molecular dynamics (MD) can directly evaluate the migration of a ligand into a protein, but they are significantly more computationally demanding~\cite{kingsley2014including}. 
Moreover, MD simulations have to be prepared for a specific ligand/protein pair, which is not practical for testing a large number of ligands.
The geometric-based methods are therefore still preferable due to their speed.

The traversability of a tunnel by a flexible ligand can be formulated as a motion planning problem and solved using sampling-based motion planners.
These planners can cope with objects (robots) with many Degrees of Freedom (DOF)  and they can also consider objects of arbitrary shape.
Two widely used planners are Probabilistic Roadmaps (PRM)~\cite{kavrakiForPP} and Rapidly Exploring Random Tree (RRT)~\cite{lavalleRRT}.
%Sampling-based planners like RRT can be used to find the trajectories even for flexible ligands, as they can cope with many-DOF systems and they can consider arbitrary shape.
%The computation of the ligand trajectories can be solved also using motion 
%In order to determine if the ligand can pass the tunnel using the geometric-based methods, it is necessary to compute a trajectory considering the shape of the ligand and its 3D translation and rotation.
%In each iteration of RRT, a random configuration $\qrand \in \C$ is generated and its nearest node $\qnear \in \T$ in the tree is found.
%A new configuration $\qnew$ is constructed on the line connecting $\qnear$ and $\qrand$ in the distance $\varepsilon$ from $\qnear$.
%If $\qnew$ is collision-free, it is added to the tree.
%The algorithm terminates if the tree approaches the goal configuration close enough.
%rrt for molecular: \cite{al2012motion}
%survey \cite{gipson2012computational}
%Solutions for tunnel detection using sampling-based methods however have not been discussed yet.
%Sampling-based planners have been used in various applications in robotics~\cite{elbanhawi2014sampling} and also  %,latombe1999motion} and also
% to study proteins, e.g. in  
%loop motions~\cite{cortes2004geometric},
%protein folding~\cite{amato2002using,raveh2009rapid,novinskaya2015improving,songPFintro},
%protein folding~\cite{raveh2009rapid,novinskaya2015improving},
%or protein folding combined with ligand diffusion~\cite{cortes2010simulating}.
The well known issue of the sampling-based planners is the narrow passage problem.
Narrow passage is a small region in the configuration space, whose removal changes the connectivity of the space.
Due to its low volume, the probability of sampling the narrow passage is low.
Consequently, many iterations are needed in order to put enough samples into the passages, which increases the computation time.
PRM-based planners can cope with the narrow passage problem by increasing the probability of sampling in difficult regions, e.g.
by sampling along medial axis~\cite{amatoOBRRT,amato2002using,wilmarthMAPRM}.
In~\cite{bergWIG}, workspace is represented by a grid in which difficult regions are identified using watershed labeling.
The regions can be estimated online, e.g., based on the number of collision-free samples around a given configuration~\cite{overmarsGauss,hsuBridge} or even identified by a human operator~\cite{denny2018general}. 
%The narrow passages are however located in the configuration space and it is generally not easy to estimate their location
%based only on the features of the workspace~\cite{hannaWIS}.
%.g. based on amount of collision-free samples around a given sample~\cite{overmarsGauss} or e.g. based
%~\cite{hsuBridge,overmarsGauss}.

@@

RRT-based planners cannot cope with the narrow passage problem simply by increasing the probability of sampling in the difficult regions, as the
growth of the configuration tree towards the random samples may be blocked by obstacles. %~\cite{vonasekphd}.
To prevent this blocking, DD-RRT~\cite{yershovaDDRRT} limits the selection of nodes for expansion to a small ball. 
The radius of the ball is set to infinity for new nodes, and decreases to a predefined radius if the node cannot be successfully expanded.
Automatic adaptation of this radius was proposed in ADD-RRT~\cite{jailletADRRT}.
%The disadvantage of DD-RRT is the sensitivity to the predefined radius.
%Therefore, the authors of ADD-RRT~\cite{jailletADRRT} proposed to adapt the ball radius according to the success rate of the expansion.
To attract the tree towards a given region, random samples have to be generated there only if the tree can expand to this region.
In~\cite{kardossRRTKK}, the probability of sampling is increased in several waypoints close to narrow passages, but without specifying
how to find these waypoints.
The generalization of~\cite{kardossRRTKK} is the guided sampling, where a whole path is used
to generate samples in the configuration space.
The path can be computed in the workspace~\cite{vonasek2009rrt}, or iteratively refined in the configuration space based on the solution of a relaxed version of the problem~\cite{bayazitIRC}.

@@

%Another techniques to sample the narrow passages dense enough is to retract the random samples towards regions that are believed
%to be located in narrow passages~\cite{zhangRetraction,lee2012srrrt}.
%the random sample $\qrand$ is connected to the tree if the line segment  from $\qnear$ to $\qrand$ is collision-free. Otherwise, the retraction-step is performed.
%The task of the retraction step is to find a contact configuration around $\qnear$ that minimizes the distance to $\qrand$.
In Retraction-based RRT~\cite{zhangRetraction}, the tree is retracted along the boundary of the obstacles in the configuration space.
This requires the analysis of the contact space, which leads to the generalized penetration depth computation of high complexity~\cite{he2016efficient}. %\cite{he2016efficient} in RAL!! + zhang2008fast
Selective Retraction-based RRT~\cite{lee2012srrrt} avoids the expensive growth in the open areas of the configuration space and focuses 
the retractions only on the narrow passages.
Obstacle-based RRT~\cite{amatoOBRRT} utilizes several expansion procedures based on, e.g., random vectors, obstacle vectors, or medial axis.
In MARRT~\cite{denny2014marrt}, the newly generated samples are pushed towards the medial axis.
%The operator can also select a suitable planner for the region.<F12>
%Therefore, many algorithms uses a heuristic to compute the samples near boundary of the configuration space~\cite{amatoOBPRM,amatoOBRRT}.

The probability of sampling of the narrow passages can be increased by dilating the free-space, e.g., by shrinking the geometry of 
the robot~\cite{hsuOnProb} or the obstacles~\cite{bayazitIRC}.
The shrinking technique has been used also in~\cite{cortes2010simulating}, where the exit pathways for a small flexible molecule are
computed using a RRT-based method.
The flexible ligand is modeled as a kinematic chain, which increases the dimension of the configuration space.
To cope with these many-DOF ligands, the RRT-ML~\cite{cortes2007mlrrt} approach is employed as the basic planner in~\cite{cortes2010simulating}.
RRT-ML expands the tree primarily using those DOF that are essential for achieving the motion of the ligand (i.e., rotation
and translation) and it employs the other DOFs (i.e., those that are responsible for conformational changes) if they hinder the growth of the tree.
In~\cite{singhLIG}, the pathways for a flexible ligand are found using PRM.
Each configuration in the roadmap is also evaluated using electrostatic energy.
The high-energy configurations are discarded, while low-energy conformations (and their vicinity) are sampled more dense.
The approach~\cite{singhLIG} can be also used to predict ligand binding sites.

>>>>

\begin{abstract}
Chemical interactions between proteins and other molecules (ligands) take place in specific locations, called active sites, that are often deeply buried inside the protein structure.
These active sites are accessible through one or more void paths, called tunnels.
Nowadays, tunnels are mostly computed using Voronoi diagrams which approximate the ligand by a bounding sphere.
This can lead to highly imprecise results.
To address this problem, we define the computation of the transportation path of non-spherical ligands including their conformation changes (i.e., when the atoms of the ligand move relatively to each other) as a motion planning task in the high-dimensional configuration space.
This can be solved using sampling-based planning methods.
We propose a modification of the Rapidly Exploring Random Tree planner for the purpose of trajectory computations for flexible ligands.
To cope with the narrow passage problem, the random samples are generated around a tunnel detected by a Voronoi-based approach.
%which serves us for better localization of the searched space.
The flexibility of the ligand is modeled using a fixed set of known conformations, which does not increase the number of dimensions of the configuration space.
The proposed method has been evaluated in a virtual screening scenario to estimate the traversability of tunnels for several ligands.
\end{abstract}

@@

The reactivity of proteins with other molecules plays a crucial role in many research disciplines, including protein engineering and drug design.
Protein reactivity is tightly connected with the presence of the void space between its amino acids. 
This space, forming a path called a tunnel, can be utilized by a small molecule (ligand) which can enter the protein inner space and react with the amino acids surrounding the protein active site~\cite{gora2013gates} (see Fig.~\ref{fig::motiv}).
The location and size of a tunnel determine the size and shape of a ligand passing through this tunnel to the active site.
Currently, the biochemists can use various computational methods for tunnel detection in order to estimate the possibility of interactions between a protein and a ligand.
This can help them to reduce the number of necessary in-vitro experiments.
The most common tools detect the tunnels using Voronoi diagrams~\cite{yaffe2008,caver3} assuming a spherical ligand (probe), 
i.e., they approximate the ligand by a bounding sphere.
However, ligands are typically of a non-spherical shape which is completely ignored by these methods.

@@

Our approach aims to overcome this problem by taking into account the actual shape of the ligand.
This necessarily requires to consider also conformational changes of the ligand, i.e., the mutual position of the ligand atoms can change.
Such a ligand is denoted as flexible ligand.
%Our approach aims to overcome this problem by taking into account the actual shape of the ligand and also its conformations.
%\textcolor{blue}{To provide the biochemists with a solution which detects tunnels also in these frames and understand the ligand passage, it is necessary to take into account not only the ligand shape but also its conformation changes when the ligand moves.}
%The proposed solution needs also to cope with the narrow passage problem, as proteins are dense structures limiting the ligand movements inside.
%Such a ligand, for which the mutual position of its atoms can change, is denoted as a flexible ligand.

The analysis of the traversability of tunnels by flexible ligands can be formulated as a motion planning problem and solved by sampling-based planners.
The sampling-based planners randomly sample the configuration space of the robot (ligand) and connect the collision-free samples into a graph structure (roadmap).
Rapidly Exploring Random Tree (RRT)~\cite{lavalleRRT} is one of the most used sampling-based planners.
Besides many applications in robotics, the sampling-based planners have been already used in biochemistry as well, 
e.g., to study loop motions~\cite{cortes2004geometric}, to identify binding sites~\cite{bayazit2001ligand}, for exit
pathway detection~\cite{cortes2010simulating,cortes2005path}, and especially for protein folding~\cite{raveh2009rapid,amato2002using}. %novinskaya2015improving

@@

In this paper, we propose a modification of the RRT planner to find trajectories for non-spherical flexible ligands inside protein tunnels.
Random samples are generated only around a given tunnel instead of the whole configuration space.
%We localize the generation of the random samples, i.e., instead of sampling the whole configuration space we focus only on the space surrounding a potential tunnel detected by the Voronoi-based method (using a smaller ligand probe in order to avoid omitting possibly feasible tunnel candidates).
The potential tunnel is represented by a set of consecutive spheres surrounding the tunnel centerline.
To enable the movement of the ligand in the narrow parts of the tunnel, the free-space is dilated by shrinking the atom radii, similarly, e.g., to~\cite{hsu06multilevel}.  %cortes2005path removed as it is not clear it they really rescale
The ligand flexibility is modeled using a fixed set of predefined conformations which can be taken by the ligand.

Finding trajectories only along a given tunnel is faster than computing all possible exit path, due to the decreased volume of the searched configuration space.
Another advantage of our method, in comparison with the existing approaches, is that we enable the users to trace the ligand path leading from the
outer solvent towards the protein active site.

The proposed method was tested by the biochemists who compared the results with already known trajectories of ligands, determined by their traditional approaches.
In the following section, we first mention several approaches and methods related to our solution and then describe the proposed method in detail.

@@

Most of the available tools for tunnel detection are based on Voronoi diagrams~\cite{yaffe2008,caver3} or they utilize a grid-based computation~\cite{sehnal2013mole}.
%The tunnels are pathways computed for a single atom (spherical probe) using Voronio-diagrams~\cite{yaffe2008,caver3} or 
%grid-based approaches~\cite{sehnal2013mole,petrek2006caver}.
Tunnels are represented as a sequence of spheres and characterized by their bottleneck, length, curvature, and the list of surrounding residues, whose physico-chemical properties are crucial for assessing the protein-ligand interaction possibilities.
More details can be found in the recently published state-of-the-art report about the analysis and visualization of biomolecular cavities~\cite{Krone_2016}.
The main disadvantage of both grid-based and VD-based methods is that the shape of the ligand is not taken into account during the tunnel detection and it is therefore not easy to estimate if (and how) a non-spherical ligand can traverse the tunnel.
The ligand flexibility cannot be principally considered by these approaches.
Simulations based on molecular dynamics (MD) can directly evaluate the migration of the ligand into the protein, but they are significantly more computationally demanding~\cite{kingsley2014including}. 
Moreover, MD simulations have to be prepared for a specific ligand/protein pair, which is not practical for testing a large number of different ligands.
The geometry-based methods are therefore still preferable due to their speed.

The traversability of a tunnel by a flexible ligand can be formulated as a motion planning problem and solved using the sampling-based motion planners.
These planners can cope with objects (robots) with many Degrees of Freedom (DOF) and they can also consider objects of an arbitrary shape.
Two widely used planners are the Probabilistic Roadmaps (PRM)~\cite{kavrakiForPP} and the Rapidly Exploring Random Tree (RRT)~\cite{lavalleRRT}.
%Sampling-based planners like RRT can be used to find the trajectories even for flexible ligands, as they can cope with many-DOF systems and they can consider arbitrary shape.
%The computation of the ligand trajectories can be solved also using motion 
%In order to determine if the ligand can pass the tunnel using the geometric-based methods, it is necessary to compute a trajectory considering the shape of the ligand and its 3D translation and rotation.
%In each iteration of RRT, a random configuration $\qrand \in \C$ is generated and its nearest node $\qnear \in \T$ in the tree is found.
%A new configuration $\qnew$ is constructed on the line connecting $\qnear$ and $\qrand$ in the distance $\varepsilon$ from $\qnear$.
%If $\qnew$ is collision-free, it is added to the tree.
%The algorithm terminates if the tree approaches the goal configuration close enough.
%rrt for molecular: \cite{al2012motion}
%survey \cite{gipson2012computational}
%Solutions for tunnel detection using sampling-based methods however have not been discussed yet.
%Sampling-based planners have been used in various applications in robotics~\cite{elbanhawi2014sampling} and also  %,latombe1999motion} and also
% to study proteins, e.g. in  
%loop motions~\cite{cortes2004geometric},
%protein folding~\cite{amato2002using,raveh2009rapid,novinskaya2015improving,songPFintro},
%protein folding~\cite{raveh2009rapid,novinskaya2015improving},
%or protein folding combined with ligand diffusion~\cite{cortes2010simulating}.
The well known issue of the sampling-based planners is the narrow passage problem.
The narrow passage is a small region in the configuration space, whose removal changes the connectivity of the space.
Due to its low volume, the probability of sampling the narrow passage is low.
Consequently, many iterations are needed in order to put enough samples into the passages, which increases the computation time.
The PRM-based planners can cope with the narrow passage problem by increasing the probability of sampling in difficult regions, e.g., by sampling along the medial axis~\cite{amatoOBRRT,amato2002using,wilmarthMAPRM}.
In~\cite{bergWIG}, the workspace is represented by a grid in which the difficult regions are identified using watershed labeling.
The regions can be estimated online, e.g., based on the number of collision-free samples around a given configuration~\cite{overmarsGauss,hsuBridge} or even identified by a human operator~\cite{denny2018general}. 
%The narrow passages are however located in the configuration space and it is generally not easy to estimate their location
%based only on the features of the workspace~\cite{hannaWIS}.
%.g. based on amount of collision-free samples around a given sample~\cite{overmarsGauss} or e.g. based
%~\cite{hsuBridge,overmarsGauss}.

@@

The RRT-based planners cannot cope with the narrow passage problem simply by increasing the probability of sampling in the difficult regions, as the growth of the configuration tree towards the random samples may be blocked by obstacles. %~\cite{vonasekphd}.
To prevent this blocking, the DD-RRT~\cite{yershovaDDRRT} limits the selection of nodes for the expansion to a small ball. 
The radius of this ball is set to infinity for new nodes, and decreases to a predefined radius if the node cannot be successfully expanded.
An automatic adaptation of this radius was proposed in the ADD-RRT~\cite{jailletADRRT}.
%The disadvantage of DD-RRT is the sensitivity to the predefined radius.
%Therefore, the authors of ADD-RRT~\cite{jailletADRRT} proposed to adapt the ball radius according to the success rate of the expansion.
To attract the tree towards a given region, random samples have to be generated there only if the tree can expand to this region.
In~\cite{kardossRRTKK}, the probability of sampling is increased in several waypoints close to narrow passages, but without specifying how to find these waypoints.
The generalization of~\cite{kardossRRTKK} is the guided sampling, where the whole path is used to generate samples in the configuration space.
The path can be computed in the workspace~\cite{vonasek2009rrt}, or iteratively refined in the configuration space based on the solution of a relaxed version of the problem~\cite{bayazitIRC}.

@@

%Another techniques to sample the narrow passages dense enough is to retract the random samples towards regions that are believed
%to be located in narrow passages~\cite{zhangRetraction,lee2012srrrt}.
%the random sample $\qrand$ is connected to the tree if the line segment  from $\qnear$ to $\qrand$ is collision-free. Otherwise, the retraction-step is performed.
%The task of the retraction step is to find a contact configuration around $\qnear$ that minimizes the distance to $\qrand$.
In the Retraction-based RRT~\cite{zhangRetraction}, the tree is retracted along the boundary of the obstacles in the configuration space.
This requires the analysis of the contact space, which leads to the generalized penetration depth computation of high complexity~\cite{he2016efficient}. %\cite{he2016efficient} in RAL!! + zhang2008fast
The Selective Retraction-based RRT~\cite{lee2012srrrt} avoids the expensive growth in the open areas of the configuration space and focuses the retractions only on the narrow passages.
The Obstacle-based RRT~\cite{amatoOBRRT} utilizes several expansion procedures based on, e.g., random vectors, obstacle vectors, or medial axis.
In MARRT~\cite{denny2014marrt}, newly generated samples are pushed towards the medial axis.
%The operator can also select a suitable planner for the region.<F12>
%Therefore, many algorithms uses a heuristic to compute the samples near boundary of the configuration space~\cite{amatoOBPRM,amatoOBRRT}.

The probability of sampling of the narrow passages can be increased by dilating the free-space, e.g., by shrinking the geometry of the robot~\cite{hsuOnProb} or the obstacles~\cite{bayazitIRC}.
The shrinking technique has been used also in~\cite{cortes2010simulating}, where the exit pathways for a small flexible molecule are computed using a RRT-based method.
The flexible ligand is modeled as a kinematic chain, which increases the dimension of the configuration space.
To cope with these many-DOF ligands, the RRT-ML~\cite{cortes2007mlrrt} approach is employed as the basic planner in~\cite{cortes2010simulating}.
RRT-ML expands the tree primarily using those DOF, that are essential for achieving the motion of the ligand (i.e., rotation
and translation) and it employs other DOFs (i.e., those that are responsible for conformational changes) if they hinder the growth of the tree.
In~\cite{singhLIG}, the pathways for a flexible ligand are found using PRM.
Each configuration in the roadmap is also evaluated using the electrostatic energy.
The high-energy configurations are discarded, while the low-energy conformations (and their vicinity) are sampled densely.
The approach~\cite{singhLIG} can be also used to predict the ligand binding sites.
